Monte Carlo tree search compared to A\(^*\) in airspace configuration decision problem

### 2018-05-18 article/article.tex
(~) Reworked results

\usepackage{fontspec}
\usepackage{polyglossia}
\usepackage{cite}
\usepackage{hyperref}
\usepackage{unicode-math}
\usepackage{amsmath,amsthm,stmaryrd}
\usepackage{algorithm,algpseudocode}
\usepackage{paralist}
\usepackage{titlesec}
\usepackage{tikz,pgfplots}

@@

\subsection{Experimental setup}
For the tests, the Paris West area is used. It contains 12 elementary modules
and 20 control sectors. The resulting model tree has a mean branching factor of
16. Since it is the smallest French area, considering the already important
branching factor, a stochastic tree search method seems appropriate.

The traffic is simulated the following way, each elementary module has an
associted list which, at index \(i\) has the number of aircraft in it at time
step \(i\). Those lists have been randomly generated. The set of lists form a
scenario which is used in all following tests.

\subsection{Accuracy}
Accuracy is assessed by comparing the exact cost given by the A\(^*\) algorithm
with the result of the MCTS\@. The time allowed to the MCTS to run is considered
long enough to allow the MCTS to converge to its best solution.

It results in figure~\ref{fig:longterm} that the algorithm described here has a
bias of approximately 15\%, considering that the stabilisation matches the
optimal value reachable.

Considering paragraph~\ref{ssec:theoretical_performances}, the MCTS should
eventually converge to the optimal value. This result reveals a flaw in the
implementation.
\begin{figure}
  \begin{tikzpicture}
    \begin{axis}
      [line width=0.05, mark size=0.1, xlabel=time(s), ylabel=cost]
      \addplot+ table {data/longterm.data};
      \addplot+[dashed, domain=0:10] {1268};
    \end{axis}
  \end{tikzpicture}
  \caption{The continuous line shows the total cost of a path computed by the
    MCTS algorithm. The x-axis shows the time allowed to the algorithm. The
    dashed curve is 15\% of the optimal cost found with the \(A^*\)
  algorithm.}\label{fig:longterm}
\end{figure}
\begin{figure}
  \begin{tikzpicture}
    \begin{axis}
      [line width=0.05, mark size=0.1, xlabel=time(s), ylabel=cost]
      \addplot+ table {data/shortterm.data};
    \end{axis}
  \end{tikzpicture}
  \caption{Monte Carlo tree search called with varying time allowed to compute
  each step, with a depth of 40 and \(\theta = 11\). The cost seems to be
  approximately stable from \(t=0.5\)s.}\label{fig:shortterm}
\end{figure}

>>>>

\usepackage{fontspec}
\usepackage{polyglossia}
\usepackage{cite}
\usepackage{hyperref}
\usepackage{unicode-math}
\usepackage{amsmath,amsthm,stmaryrd}
\usepackage{algorithm,algpseudocode}
\usepackage{paralist}
\usepackage{titlesec}
\usepackage{tikz,pgfplots}
\usepackage{array,tabulary,booktabs}

@@

\subsection{Experimental setup}
For the tests, a fictitious area is used. It contains 5 elementary modules
and 12 control sectors. The resulting model tree has a mean branching factor of
6.
Regarding the branching, the smallest French area -- which is Paris West
-- has a branching factor of 16 for 20 control sectors and 12 elementary
modules. Considering the already important
branching factor, a stochastic tree search method seems appropriate.

The traffic is simulated the following way, each elementary module has an
associted list which, at index \(i\) has the number of aircraft in it at time
step \(i\). Those lists have been written by hand, the activity shows a traffic
peak at the beginning of the scenario and much less activity after. The expected
outcome of the algorithms is a massive degrouping reflecting the heavy workload
followed by merges.

\subsection{Accuracy}
Accuracy is assessed by comparing the outcome of the MCTS algorithm with the
A\(^*\) cost and a greedy algorithm cost.
The time allowed to the MCTS to run is considered long enough to allow the MCTS
to converge to its best solution. The results are summarised in
table~\ref{tab:methods_compare}.

The graph~\ref{fig:grouping} asserts the correct behaviour of both algorithms
A\(^*\) and MCTS regarding grouping and degrouping.

Considering paragraph~\ref{ssec:theoretical_performances}, the MCTS should
eventually converge to the optimal value. This result reveals a flaw in the
implementation.

\begin{table}
  \centering
  \begin{tabulary}
    {\textwidth}{LRR}
    Alg. & Cost & Err.\\\toprule
    A\(^*\) & 77.31 & 0\%    \\
    MCTS    & 82.48 & 6.7\%  \\
    Greedy  & 90.50 & 17.1\% \\
  \end{tabulary}
  \caption{Comparison of several methods on the generated
  scenario.}\label{tab:methods_compare}
\end{table}

\begin{figure}
  \begin{tikzpicture}
    \begin{axis}
      [line width=0.05, mark size=0.1, xlabel=timestep, ylabel=num of sectors,
      legend entries={A\(^*\), MCTS}]
      \foreach \i in {0,1}{%
        \addplot+ table [
          x expr={\lineno}, y index=\i
          ]
          {./data/grouping.data};
      }
    \end{axis}
  \end{tikzpicture}
  \caption{Grouping profile}\label{fig:grouping}
\end{figure}
\begin{figure}
  \begin{tikzpicture}
    \begin{axis}
      [line width=0.05, mark size=0.1, xlabel=time(s), ylabel=cost]
      \addplot+ table {data/longterm.data};
      \addplot+[dashed, domain=0:10] {1268};
    \end{axis}
  \end{tikzpicture}
  \caption{The continuous line shows the total cost of a path computed by the
    MCTS algorithm. The x-axis shows the time allowed to the algorithm. The
    dashed curve is 15\% of the optimal cost found with the \(A^*\)
  algorithm.}\label{fig:longterm}
\end{figure}
\begin{figure}
  \begin{tikzpicture}
    \begin{axis}
      [line width=0.05, mark size=0.1, xlabel=time(s), ylabel=cost]
      \addplot+ table {data/shortterm.data};
    \end{axis}
  \end{tikzpicture}
  \caption{Monte Carlo tree search called with varying time allowed to compute
  each step, with a depth of 40 and \(\theta = 11\). The cost seems to be
  approximately stable from \(t=0.5\)s.}\label{fig:shortterm}
\end{figure}
